Report on the Third Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE3)

### 2016-08-30 WSSSPE3/report/WSSSPE3_review.tex
adding RSE cites/references

\item \textbf{James Hetherington, Jonathan Cooper, Robert Haines, Simon
Hettrick, James Spencer, Mark Stillwell, Mike Croucher, Christopher Woods, and
Susheel Varma: \textit{An update from UK Research Software Engineers}}. James
Hetherington started by listing some of the problems research software faces,
which include poor standard of verification and low levels of reuse. For a long
time, technical solutions to such technical problems were focus of the eResearch
community, including research software distributions, grids, middleware and
workflows. Some limited adoption can be seen today in research communities, but
the main problems have not been solved to a sufficient level. Hetherington
hypothesized that instead of technical solutions, social innovation is needed: a
new role in the academic system focused on research software that combines the
best parts of a craftsperson and a scholar. However, social innovation in
centuries-old institutions is hard. Alternatives to such a new role would have
to include rewards for good research software, recognition of software as
academic output, and rejecting submissions based on irreproducible computational
results. Some advantages of research software engineering (RSE) groups include
the possibility of training in reproducible computational research, providing
collaborations for researchers who do not want to serve as programmers, and
creating synergies with research computing platforms. The success of such a
group would be measured by the output and quality of the research software.
Members could be part of Research Computing or faculty. They would not be
independent researchers, but would have to have a research background. An
attempt to form a community of RSE groups within the UK has been underway for
several years, including funding from the UK Research Council. An open question
is whether this approach can be adopted in other countries, including the~USA.

>>>>

\item \textbf{James Hetherington, Jonathan Cooper, Robert Haines, Simon
Hettrick, James Spencer, Mark Stillwell, Mike Croucher, Christopher Woods, and
Susheel Varma: \textit{An update from UK Research Software Engineers}}. James
Hetherington started by listing some of the problems research software faces,
which include poor standard of verification and low levels of reuse. For a long
time, technical solutions to such technical problems were focus of the eResearch
community, including research software distributions, grids, middleware and
workflows. Some limited adoption can be seen today in research communities, but
the main problems have not been solved to a sufficient level. Hetherington
hypothesized that instead of technical solutions, social innovation is needed: a
new role in the academic system focused on research software that combines the
best parts of a craftsperson and a scholar~\cite{rse1,rse2}. However, social innovation in
centuries-old institutions is hard. Alternatives to such a new role would have
to include rewards for good research software, recognition of software as
academic output, and rejecting submissions based on irreproducible computational
results. Some advantages of research software engineering (RSE) groups include
the possibility of training in reproducible computational research, providing
collaborations for researchers who do not want to serve as programmers, and
creating synergies with research computing platforms. The success of such a
group would be measured by the output and quality of the research software.
Members could be part of Research Computing or faculty. They would not be
independent researchers, but would have to have a research background. An
attempt to form a community of RSE groups within the UK has been underway for
several years, including funding from the UK Research Council. An open question
is whether this approach can be adopted in other countries, including the~USA.
